---
title: "cross-cafe-ARIMA"
author: "Timon Florian Godt"
date: "5/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# set working directory
setwd("~/BI-2020/Data-Science/Projects/Data-Science-Project/Data")
```

```{r}
# import libraries
pacman::p_load(ISLR, dplyr, ggplot2, tidyverse, GGally, corrplot, caret, 
               e1071, MASS, class, readxl, reshape2, keras, scales, gridExtra,
               grid, zoo, lubridate, forecast, tseries, timetk, tidyquant,
               tibbletime, forcats, glue, openair, stats)
```

```{r}
# load dataset
cross <- read_csv("cross-cafe-final.csv")
spec(cross)
```

```{r}
# isolate cross cafe sales data
cross <- cross[, c(2, 4)]
```

```{r}
# isolate lockdown period
names(cross)[1] <- "date"
cross2020 <- selectByDate(
  cross,
  start = "2020-05-18",
  end = "2020-12-08 ",
)
```

```{r}
# transform cross cafe into time series 
firstHour.2020<- 24*(as.Date("2020-5-18 00:00:00")-as.Date("2020-12-18 00:00:00"))
cc.ts <- ts(cross2020$cross_cafe, start(2020,firstHour.2020), frequency = 24)
# plot cross cafe sales data
ggplot(cross2020, aes(x = date, y = cross_cafe)) + geom_line()
```

```{r}
ggplot(cross2020[1:240,], aes(x = 1:240, y = cross_cafe)) + geom_line()
```
the data seems seasonal in the sense that sales only occur during the day. it also appears that some days (most likely the weekend) have higher sales than others.
```{r}
# check for autocorrelation for the last 10 days
acf(cc.ts, lag.max = 240)
```
Autocorrelation is pretty high, which is good for our forecasting. All autocorrelation values are inside confidence bounds. It seems as if the data is highly correlated which would explain why
a simple naive approache could be effective. However, a ARIMA model is very good at picking up on this correlation. Let's see where we go from here.

```{r}
pacf(cc.ts, lag.max = 240)
```

```{r}
# inspect for stationarity 
adf.test(cc.ts)
```
The p-value is very small, therefore suggesting that we our series is stationary.
```{r}
# decompose time series to isolate a potential trend or seasonal component
plot(decompose(cc.ts, type = "additive"))
```

```{r}
plot(decompose(cc.ts, type = "multiplicative"))
```

```{r}
# split data into train and test set
size_train = round(length(cc.ts) - 240)
train_data <- ts(cc.ts[1:size_train], start = 1)
test_data <- ts(cc.ts[(size_train +1):length(cc.ts)], start = size_train +1)
```

```{r}
# fit auto.arima on whole series to find out the optimal specification
cc.fit <- auto.arima(cc.ts, max.p = 240, max.q = 240, seasonal = TRUE)
summary(cc.fit)
# Series: cc.ts 
# ARIMA(2,0,0)(2,1,0)[24] 

```

```{r}
# fit model on train data
fit.train <- Arima(train_data, order = c(2,0,0), seasonal = c(2,1,0)[24])
summary(fit.train)
# checking model diagnostics
tsdisplay(residuals(fit.train), main='Model Residuals')
```

```{r}
# Arima forecast for cross cafe sales
arima.fcast <- forecast(fit.train, h= 240)
plot(arima.fcast)
lines(cc.ts)

accuracy(arima.fcast, train_data)
```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```